---
title: "Lab 12 - Smoking during pregnancy"
author: "Steph Bossert"
date: "22 April 21"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(viridis)
library(openintro)
data(ncbirths)
view(ncbirths)
?ncbirths
# set a seed
set.seed(1234)
```

### Exercise 1
```{r EX1}
###Cases= Random sample of 1000 births in NC in 2004; there are 1000 cases
ggplot(data = ncbirths, aes(x= visits, fill=habit)) +
  geom_bar(stat = "count", position = position_dodge())
ggplot(data = ncbirths, aes(x= lowbirthweight, fill=habit)) +
  geom_bar(stat = "count", position = position_dodge())
ggplot(data = ncbirths, aes(x= premie, fill=habit)) +
  geom_bar(stat = "count", position = position_dodge())
ggplot(data = ncbirths, aes(x= visits, fill=whitemom)) +
  geom_bar(stat = "count", position = position_dodge())
###Outliers- visits 


```
### Exercise 3
###Mean = 7.25
###Exercise 4: Yes, conditions are satisfied because 7.25 /=/ 7.43, this is the estimate of population value. Our sample mean does not equal that point estimate, so we should conduct additional analysis. 
```{r Ex3}
ncbirths_white <- ncbirths %>%
  filter(whitemom == "white")
view(ncbirths_white)
summary(ncbirths_white$weight)

```
###Exercise 5
```{r Bootstrap}
boot_ncw <- ncbirths_white %>%
  # specify the variable of interest
  specify(response = weight) %>% 
  # generate 15000 bootstrap samples
  generate(reps = 15000, type = "bootstrap")%>% 
  # calculate the mean of each bootstrap sample
  calculate(stat = "mean")

#X= 7.25 - 7.251 = .09

ggplot(data = boot_ncw, mapping = aes(x = stat)) +
  geom_histogram()+#binwidth = 2
  labs(title = "Bootstrap distribution of means")
###create new graph with shifted center of distro
#shift this distribution to be centered at the null value by subtracting / adding X to all boostrap mean (X = difference between mean of bootstrap #distribution and null value), and
#calculate the p-value as the proportion of bootstrap samples that yielded a sample mean at least as extreme as the observed sample mean.

boot_ncw2 <- boot_ncw %>%
  mutate(newmean = stat + .09)
view(boot_ncw2)
ggplot(data = boot_ncw2, mapping = aes(x = newmean)) +
  geom_histogram()+#binwidth = 2
  labs(title = "Bootstrap distribution of null means")
#YAY!
###p< .05- it appears the weights have changed since 1995. 

```
 

```{r Baby weight}
ncbirths %>%
  ggplot( aes(x=weight, y=habit, fill=habit)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=9, color="red", fill="red") +
  coord_flip() +
    ggtitle("Baby Weight by Mother's Smoking Habit") +
    xlab("")
```

```{r Ex 7-10}
ncbirths_habitgiven <- ncbirths %>%
  filter(habit != "NA")
ncbirths_habitgiven %>%
  group_by(habit)
view(ncbirths_habitgiven)

boot_ncw <- ncbirths_habitgiven %>%
  # specify the variable of interest
  specify(response = weight) %>% 
  # generate 15000 bootstrap samples
  generate(reps = 15000, type = "bootstrap") %>% 
  # calculate the mean of each bootstrap sample
  calculate(stat = "mean")
#view(boot_ncw)
res.ftest <- var.test(weight ~ habit, data = ncbirths_habitgiven)
res.ftest

ggplot(data = boot_ncw, mapping = aes(x = stat)) +
  geom_histogram()+#binwidth = 2
  labs(title = "Bootstrap distribution of means")
ncbirths_habitgiven %>%
  specify(response = weight) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "mean") %>%
  summarize(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))
res.ftest <- var.test(weight ~ habit, data = ncbirths_habitgiven)
res.ftest
####p = .1989
### Null: Weights of babies from smoking mothers is the same as the weight of babies from nonsmoking mothers
### H0: Weight(baby_smoke) = Weight(baby_nonsmoke)
###p value indicates the two groups do NOT differ signficantly in baby weight 
```
### The baby weights from mom's who smoke and those who do not are not signficiantly different 
###Ex 8: yes conditions are necessary to run additional tests because the mean weights of the two groups are not equal (which is what the null predicts) 
###Ex 10: P value indicates that the weights of the babies from the two groups are not significantly different 
###Ex 11 Age cutoff is 35- create new datframe with only mature moms, then run summary for range of ages
```{r Ex 11}
age <- ncbirths %>%
  filter(mature == "mature mom")
view(age)
summary(age$mage)
```

```{r Ex 12}
maturebirth <- ncbirths[,c(3,10)]
view(maturebirth)
maturebirth %>%
  count(mature, lowbirthweight)
maturebirth %>%
  count(mature,lowbirthweight) %>%
  group_by(mature) %>%
  mutate(p_hat = n / sum(n))
# mature mom: 18/133= .14

# younger mom: 93/867 = .11
null_dist <-maturebirth %>%
 specify(response = lowbirthweight, explanatory = mature, 
          success = "not low") %>%
 hypothesize(null = "independence") %>%
generate(100, type = "permute") %>%
  specify(response = lowbirthweight, explanatory = mature, 
          success = "not low") %>%
calculate(stat = "diff in props", 
           order = c("mature mom", "younger mom"))
ggplot(data = null_dist, mapping = aes(x = stat)) +
  geom_histogram() +#binwidth = 0.2 
  labs(title = "Null distribution")
#this graph will not work...R keeps crashing....
null_dist %>%
  filter(stat >= 0.03) %>%
  summarize(p_value = n()/nrow(null_dist))
###Since the p value is less than .05 it appears that the proportion of low birth weight babies is higher for mature mothers 
```
